### 2017-07-15 draft/unique-linear.tex
polymorphic-woes

  \begin{array}{|c|c||c|c||c|c|l}
    Q_a&Q_b&O_a&O_b&P_a&P_b& \\
    \hline
    P_a\sqcup\READ &\WRITE& \NOTHING & \WRITE & P_a & \WRITE
                           & \text{copy, propagate left to write} \\
    \READ & \WRITE& O_a & \WRITE & \NOTHING & \WRITE
                           & \text{copy, propagate left to write} \\
    P_a\sqcup\READ   &\WRITE& \READ & \WRITE & P_a & \WRITE
                           &  \text{copy, propagate left to write}\\
    \READ&\WRITE& \WRITE & \WRITE & \READ & \WRITE
                       &  \text{copy, propagate left to write}\\
    \WRITE &\WRITE& \WRITE & \WRITE & \WRITE & \WRITE
                       &\text{copy, propagate left to write} \\
  \end{array}
\end{mathpar}
A copy operation reads \textbf{every} field. Hence, a copy operation
caused by one field implies a read effect on all fields.
\clearpage{}
Consider the typing for 
\begin{align*}
  f
  &=(\LAM r{\REC{ a:^\WRITE \NUM}} \UPD r a{42})
  &&: \REC{ a:^\WRITE \NUM} \to \REC{ a:^\PVAR \NUM}
  \\
  g
  &=(\LAM r{\REC{a :^\NOTHING \NUM}} \REC{a = 0})
  &&: \REC{a :^\NOTHING \NUM} \to \REC{ a:^\PVAR \NUM}
  \\
  &\IF e{f}{g}
  &&: \REC{ a:^\WRITE \NUM} \to \REC{ a:^\PVAR \NUM}
\end{align*}
This propagation is correct,
\textbf{but} if the write is propagated, this may cause a useless copy if $e$ is false.
\end{document}

>>>>

  \begin{array}{|c|c||c|c||c|c|l}
    Q_a&Q_b&O_a&O_b&P_a&P_b& \\
    \hline
    P_a\sqcup\READ &\WRITE& \NOTHING & \WRITE & P_a & \WRITE
                           & \text{copy, propagate left to write} \\
    \READ & \WRITE& O_a & \WRITE & \NOTHING & \WRITE
                           & \text{copy, propagate left to write} \\
    P_a\sqcup\READ   &\WRITE& \READ & \WRITE & P_a & \WRITE
                           &  \text{copy, propagate left to write}\\
    \READ&\WRITE& \WRITE & \WRITE & \READ & \WRITE
                       &  \text{copy, propagate left to write}\\
    \WRITE &\WRITE& \WRITE & \WRITE & \WRITE & \WRITE
                       &\text{copy, propagate left to write} \\
  \end{array}
\end{mathpar}
A copy operation reads \textbf{every} field. Hence, a copy operation
caused by one field implies a read effect on all fields.
\clearpage{}
Consider the typing for 
\begin{align*}
  f
  &=(\LAM r{\REC{ a:^\WRITE \NUM}} \UPD r a{42})
  &&: \REC{ a:^\WRITE \NUM} \to \REC{ a:^\PVAR \NUM}
  \\
  g
  &=(\LAM r{\REC{a :^\NOTHING \NUM}} \REC{a = 0})
  &&: \REC{a :^\NOTHING \NUM} \to \REC{ a:^\PVAR \NUM}
  \\
  &\IF e{f}{g}
  &&: \REC{ a:^\WRITE \NUM} \to \REC{ a:^\PVAR \NUM}
\end{align*}
This propagation using $\WRITE\sqcup\NOTHING = \WRITE$ is correct,
\textbf{but} the propagation of $\WRITE$  may cause a useless copy if $e$ is false.

The typing is analogous to pushing the conditional into the lamdas.
\begin{align*}
  r : \REC{ a:^\WRITE \NUM} & \vdash \UPD r a{42} &&: \REC{ a:^\PVAR \NUM}
  \\
  r : \REC{ a:^\NOTHING \NUM} & \vdash \REC{a = 0} &&: \REC{ a:^\PVAR \NUM}
  \\
  r : \REC{ a:^\WRITE \NUM}& \vdash \IF e{( \UPD r a{42})}{(\REC{a = 0})} &&: \REC{ a:^\PVAR \NUM}
  \\
  \vdash \LAM r {\REC{ a:^\WRITE \NUM}} & \IF e{( \UPD r a{42})}{(\REC{a = 0})}
                                                  &&: {\REC{ a:^\WRITE \NUM}} \to \REC{ a:^\PVAR \NUM}
\end{align*}

\textbf{Let's finally get down to polymorphism. }
\begin{align*}
  \mathit{id} =
  \LAM x\TVAR x & : \forall\TVAR.\TVAR\to\TVAR
  &&\text{seems fine, just instantiate}
  \\
  \mathit{dup} =
  \LAM x\TVAR x & : \forall\TVAR.\TVAR\to(\TVAR,\TVAR)
\end{align*}
\begin{verbatim}
lambda r.
  let (r1, r2) = dup r 
  in  (r1 { a = 42 }, r2.a)
-- should behave like "write before read"
lambda r. 
      (r  { a = 42 }, r .a)
\end{verbatim}
\begin{align*}
  r : \REC{ a:^\WRITE \NUM} & \vdash \UPD r a{42} &&: \REC{ a:^\PVAR \NUM}
  \\
  r : \REC{ a:^\READ \NUM} & \vdash  r. a &&: \NUM
  \\
  r : \REC{ a:^\READ \NUM} & \vdash (\UPD r a{42}, r. a)  &&: \REC{ a:^\PVAR \NUM} \times \NUM
\end{align*}
Using \textit{dup}
\begin{align*}
  r_1 : \REC{ a:^\WRITE \NUM} & \vdash \UPD r a{42} &&: \REC{ a:^\PVAR \NUM}
  \\
  r_2 : \REC{ a:^\READ \NUM} & \vdash  r. a &&: \NUM
  \\
  r_1 : \REC{ a:^\WRITE \NUM},
  r_2 : \REC{ a:^\READ \NUM} & \vdash  (\UPD r a{42} , r. a) &&: \REC{ a:^\PVAR \NUM} \times \NUM
\end{align*}
\end{document}
